Optimal Fitting of Planar Curves to Prescribed Constraints

### 2013-08-15 idp/documentation/paper.tex
some reorganization, started working on solution section

	\section{Usability of Curve Design Tools}
	\label{section:usability_curve_design_tools}

		When trying to improve something, it seems advisable to take a step back and thoroughly research the causes of its perceived shortcomings so as to find a way to improve it without making the same or similar mistakes. In the context of curve design tools, this can be achieved by analyzing and modelling the process that people use when designing curves. Once such a model has been obtained, it can be used to propose criteria for good curve design software as well as to analyze potential issues with existing tools.

@@

			By observing the process that most people seem to go through when designing curves using vector graphics software, the following model appears plausible.

@@

		\subsection{Usability Analysis}
		\label{section:usability_analysis}

@@

			\subsubsection{Common Design Tools for Bézier Splines}
			\label{section:common_design_tools_bézier_splines}

@@

				Applying the criteria established in section \ref{section:usability_analysis}, we conclude that the design tools for Bézier splines work sufficiently well in terms of humans trying to read and/or speak their specification language (criteria 2 and 3), a skill that can be acquired with some practice. Bézier splines are furthermore very simple to derive from their specification, fulfilling the 5th criterion perfectly. Design tools for Bézier splines have some issues with both the fairness measure (criterion 4) and the expressiveness of the specification language (criterion 1), both of which shall be analyzed in the following paragraphs.

@@

			\subsubsection{Common Design Tools for Spiro Splines}

@@

				We analyze Spiro splines in terms of the criteria presented in section \ref{section:usability_analysis}. The specification language of Spiro splines being simply points on the curve, plus the fact that on sufficiently smooth curves, not very many of them being needed, the language is well-suited to be read and spoken by humans (criteria 2 and 3). The fairness measure leads to beautifully smooth curves that are ususally sufficiently close to what a human would expect given the control points (criterion 4). The software is also good and sufficiently quick at constructing these curves (there are some cases where the algorithm fails to find a meaningful curve, but these cases are fairly artificial and can usually be avoided when designing actual curves) and thus fulfills criterion 5 well enough.

				The only real issue lies in the insufficient expressiveness of the specification language. Section \ref{section:common_design_tools_bézier_splines} discussed how the lack of ability to specify curvature using common Bézier spline design tools adversively affects usability. Many of the same points apply to Spiro spline design tools, since they allow the user to neither specify direction nor curvature. These properties can only be expressed indirectly, by placing the control points in such a way that the algorithm will create a curve that has the desired properties, resulting in the usability issues described previously, such as requiring many control points to be specified even though the curve could be described much more concisely in terms of direction and/or curvature. This seems especially unfortunate when considering that the spline primitive of Spiro splines, the Euler spiral, would lend itsself excellently to describing circle arcs, spirals and other shapes exhibiting constant or linearly changing curvature. Sometimes, the algorithm can be coaxed into producing the desired curve by placing just a few control points very carefully, resulting in a series of Euler spiral segments that exhibit the wanted properties. Unfortunately, this is rarely practical since it both requires a lot of time to set up and doesn't behave very well when modified. These issues are somewhat alleviated by the guaranteed \(G^2\) continuity together with the general tendency of the algorithm to produce extremely smooth curves even when many control points are involved. That way, when indirectly specifying direction and/or curvature by placing many control points, it's not as easy to accidently create small bumps as with Bézier splines.

@@

				Example 1 shows a shape that is in principle well-suited for being represented by parts of the Euler spiral. Unfortunately, describing the curve using Spiro splines still requires many control points. In contrast to the corresponding example in section \ref{section:common_design_tools_bézier_splines}, the curve in this example is close to a perfect spiral an thus does not match the source image perfectly. Additional control points would have to be added in order to approximate the imperfect spiral in the source image better.

@@

@@

			We think that in order to get the best usability, using a good specification language which does not rely on implementation details has to be the top priority. Since it is not immediately clear which specification language is the best, the first step is to move away from the traditional explicit curve construction, and instead use a process that enables flexible creation of curves based on some specification and some fairness measure, as illustrated in figure \ref{figure:specification-based_curves}.

@@

	\section{Proposed Solution: Nonlinear Optimization on Polynomial Curves}

		% instance of specification-based curves
		% segmentation, separation of segmentation and specification, global positions on curve (https://code.google.com/p/kurve/wiki/SegmentationParametrization, maybe diagram)
		% disambiguation
		% specifications
		%   differential geometric properties
		%   position, direction and curvature
		% continuity connections
		% fixed length, constant velocity
		%   fixed length is necessary if specification positions are a fraction of the arc-length of the whole curve, otherwise, any change in the specifications may change the curve length and thus move all specifications along the curve
		% fairness (quote paper or thesis on MVC)
		%   derivative of projected acceleration is equal to projected jerk

@@

		\begin{table}[htbp]
			\centering
			\begin{tabular}{c|c|r|r|}
				\multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{2}{c}{Actual Choice} \\
				\cline{2-4}
				& Payout & \multicolumn{1}{|c|}{A and B} & \multicolumn{1}{|c|}{B only} \\
				\cline{2-4}
				\multirow{2}{*}{Predicted Choice} & A and B & 1 000 \$ & 0 \$ \\
				\cline{2-4}
				& B only & 1 001 000 \$ & 1 000 000 \$ \\
				\cline{2-4}
			\end{tabular}
			\caption{Possible payouts for Newcomb's paradox}
			\label{table:NewcombParadoxPayouts}
		\end{table}

>>>>

	\section{Problem Analysis}
	\label{section:problem_analysis}

		When trying to improve something, it seems advisable to take a step back and thoroughly research the causes of its perceived shortcomings so as to find a way to improve it without making the same or similar mistakes. In the context of curve design tools, this can be achieved by analyzing and modelling the process that people use when designing curves. Once such a model has been obtained, it can be used to propose criteria for good curve design software as well as to analyze potential usability issues with existing tools.

@@

			Observing the process that most people seem to go through when designing curves using vector graphics software, the following model appears plausible.

@@

		\subsection{Usability Criteria for Curve Design Tools}
		\label{section:usability_criteria_curve_design_tools}

@@

			\subsubsection{Bézier Splines}
			\label{section:bézier_splines}

@@

				Applying the criteria established in section \ref{section:usability_criteria_curve_design_tools}, we conclude that the design tools for Bézier splines work sufficiently well in terms of humans trying to read and/or speak their specification language (criteria 2 and 3), a skill that can be acquired with some practice. Bézier splines are furthermore very simple to derive from their specification, fulfilling the 5th criterion perfectly. Design tools for Bézier splines have some issues with both the fairness measure (criterion 4) and the expressiveness of the specification language (criterion 1), both of which shall be analyzed in the following paragraphs.

@@

			\subsubsection{Spiro Splines}
			\label{section:spiro_splines}

@@

				We analyze Spiro splines in terms of the criteria presented in section \ref{section:usability_criteria_curve_design_tools}. The specification language of Spiro splines being simply points on the curve, plus the fact that on sufficiently smooth curves, not very many of them being needed, the language is well-suited to be read and spoken by humans (criteria 2 and 3). The fairness measure leads to beautifully smooth curves that are ususally sufficiently close to what a human would expect given the control points (criterion 4). The software is also good and sufficiently quick at constructing these curves (there are some cases where the algorithm fails to find a meaningful curve, but these cases are fairly artificial and can usually be avoided when designing actual curves) and thus fulfills criterion 5 well enough.

				The only real issue lies in the insufficient expressiveness of the specification language. Section \ref{section:bézier_splines} discussed how the lack of ability to specify curvature using common Bézier spline design tools adversively affects usability. Many of the same points apply to Spiro spline design tools, since they allow the user to neither specify direction nor curvature. These properties can only be expressed indirectly, by placing the control points in such a way that the algorithm will create a curve that has the desired properties, resulting in the usability issues described previously, such as requiring many control points to be specified even though the curve could be described much more concisely in terms of direction and/or curvature. This seems especially unfortunate when considering that the spline primitive of Spiro splines, the Euler spiral, would lend itsself excellently to describing circle arcs, spirals and other shapes exhibiting constant or linearly changing curvature. Sometimes, the algorithm can be coaxed into producing the desired curve by placing just a few control points very carefully, resulting in a series of Euler spiral segments that exhibit the wanted properties. Unfortunately, this is rarely practical since it both requires a lot of time to set up and doesn't behave very well when modified. These issues are somewhat alleviated by the guaranteed \(G^2\) continuity together with the general tendency of the algorithm to produce extremely smooth curves even when many control points are involved. That way, when indirectly specifying direction and/or curvature by placing many control points, it's not as easy to accidently create small bumps as with Bézier splines.

@@

				Example 1 shows a shape that is in principle well-suited for being represented by parts of the Euler spiral. Unfortunately, describing the curve using Spiro splines still requires many control points. In contrast to the corresponding example in section \ref{section:bézier_splines}, the curve in this example is close to a perfect spiral an thus does not match the source image perfectly. Additional control points would have to be added in order to approximate the imperfect spiral in the source image better.

@@

	\section{Proposed Solution}
	\label{section:proposed_solution}

		TODO

@@

			We think that in order to get the best usability, using a good specification language which does not rely on implementation details has to be the top priority. Since it is not immediately clear which specification language is the best, the first step is to move away from the traditional explicit curve construction, and instead use a process that enables the creation of curves based on some specification and some fairness measure.

@@

			Figure \ref{figure:specification-based_curves} shows such a process. Note that many specifics are still left undetermined in this approach since the previous usability analysis didn't give a complete answer as to what the optimal curve design tool looks like. It has yet to be determined what the specification language and fairness measure should be, whether to do segmentation (spline approach) or use a single parametric curve, or if the curve should be determined by numerical optimization or by explicit construction. These choices will be dealt with in the following section.

		\subsection{Nonlinear Optimization on Polynomial Curves}
		\label{section:nonlinear_optimization_polynomial_curves}

			After having established the general approach to curve design tools that the usability analysis seems to suggest in section \ref{section:specification-based_curves}, it's time to decide on the specifics in order to create an instance of this model.

			The goal of this project is to figure out which specification language and fairness measure yield the best usability in the resulting curve design tool. In order to achieve this goal, we decided to use nonlinear optimization on polynomial splines, that is, the curves are splines built from parametric polynomial curves which are connected with \(G^2\) continuity and whose parameters are determined using nonlinear optimization on some objective function determined by the specification and the fairness measure. This approach allows us to conveniently experiment with different specification languages and fairness measures in order to assess their usability.

			\begin{table}[htb]
				\centering
				\begin{tabular}{c|c|r|r|}
					\multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{2}{c}{Segmentation Visibility} \\
					\cline{2-4}
					& Scenario & \multicolumn{1}{|c|}{Exposed} & \multicolumn{1}{|c|}{Opaque} \\
					\cline{2-4}
					\multirow{2}{*}{Position Specification} & Default & 1 & 2 \\
					\cline{2-4}
					& Arc-Length & 3 & 4 \\
					\cline{2-4}
				\end{tabular}
				\caption{Choices for Position Specification and Segmentation Visibility}
				\label{table:choices_position_specification_segmentation_visibility}
			\end{table}

		% segmentation, separation of segmentation and specification, global positions on curve (https://code.google.com/p/kurve/wiki/SegmentationParametrization, maybe diagram)
		% disambiguation
		% specifications
		%   differential geometric properties
		%   position, direction and curvature
		% continuity connections
		% fixed length, constant velocity
		%   fixed length is necessary if specification positions are a fraction of the arc-length of the whole curve, otherwise, any change in the specifications may change the curve length and thus move all specifications along the curve
		% fairness (quote paper or thesis on MVC)
		%   derivative of projected acceleration is equal to projected jerk

@@
